---
title: "Analysis supporting ICTMC2019 submission"
author: "Kristian Brock"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

Load data:
```{r setup, message=FALSE}
library(glue)
library(ggplot2)
library(broom)
library(purrr)
library(broom)
library(tidyr)

library(brms)
library(tidybayes)
library(ggridges)

source('../../Load.R')
```

Dose finding trials investigate many doses of an experimental treatment or combination.

# Fit separate GLMs to each series
Having collected data, we estimate the probabilitiy of DLT and objective response on dose-level.

#### DLTs
Our outcome for analysing general toxicity in these trials is dose-limiting toxicity, or *DLT*.

Our models will be very simple.
Each takes an intercept term and a gradient with respect to dose-level, using the logit link function.
Thus, average toxicity is estimated via the intercept term, and the average sensitivity to dose increases is estimated via the gradient term.
No further complexity is considered given the general tiny sample size of dose-finding trials.
```{r, warning=FALSE}
dlt %>% 
  group_by(Study, AnalysisSeriesId) %>% 
  nest() %>% 
  mutate(
    num_dose = map_int(data, nrow),
    num_pat = map_dbl(data, .f = function(df) sum(df$n)),
    num_event = map_dbl(data, .f = function(df) sum(df$Events)),
    model = map(.f = function(df) {
      glm(ProbEvent ~ 1 + DoseLevel, data = df, weights = n,
          family = binomial('logit')) }, 
      .x = data),
    tidy_model = map(model, tidy),
    fitted = map(model, augment, type.predict = 'response'),
    # Column 2, row 2 of tidy model is estimated slope parameter for dose-level
    dose_grad = map(tidy_model, 2) %>% map_dbl(2) 
  ) -> dlt_models
```

Many of these models report numerical problems with probabilities of 0 or 1 being fit.
Once again, the small average sample size is a problem.
For instance:

```{r}
dlt_models$num_pat %>% summary()
```
The median number of patients in a model is 20; the mean is 22.7.

```{r}
dlt_models$num_event %>% summary()
```
The median number of DLTs seen is 3.

Thus we are trying to model rare events with small sample sizes. 
The *separate models* approach is very unlikely to provide a robust analysis. 
A hierarchical bayesian analysis may improve inference dramatically if the various weak insights from many models can be combined in a sensible and robust way.
We consider that below.
For now, we continue with the GLM analysis to see what the data reveal.


```{r}
dlt_models$num_dose %>% summary()
```
The average number of doses is  4.

#### Objective response
We use objective response (either by RECIST or any of the disease-specific analogues like the Cheson criteria) to indicate efficacy because these are the outcomes that are most commonly reported in trials when efficacy is assessed.

Fitting the equivalent models that we fit before:
```{r, warning=FALSE}
obj_resp %>% 
  group_by(Study, AnalysisSeriesId) %>% 
  nest() %>% 
  mutate(
    num_dose = map_int(data, nrow),
    num_pat = map_dbl(data, .f = function(df) sum(df$n)),
    num_event = map_dbl(data, .f = function(df) sum(df$Events)),
    model = map(.f = function(df) {
      glm(ProbEvent ~ 1 + DoseLevel, data = df, weights = n,
          family = binomial('logit')) }, 
      .x = data),
    tidy_model = map(model, tidy),
    fitted = map(model, augment, type.predict = 'response'),
    # Column 2, row 2 of tidy model is estimated slope parameter for dose-level
    dose_grad = map(tidy_model, 2) %>% map_dbl(2) 
  ) -> obj_resp_models
```


#### Inferences
The sign of the estimated sensitivities to `DoseLevel` provide information on the tendency for binary toxicity and efficacy outcomes to become more likely as dose is increased:
```{r}
mean(dlt_models$dose_grad > 0)
mean(obj_resp_models$dose_grad > 0)
```

In 78% of the fitted DLT models, DLT incidence is seen to be positively related to dose level.
This is consistent with the adage _the dose makes the poison_, i.e. higher doses are more likely to induce adverse events.
In constrast, half the models suggest a positive association between dose and efficacy, implying that the other half suggest the opposite.

It may be illuminating to pair up the toxicity and efficacy signals within trial.
We may be interested in whether positive association with dose is coincident in the two outcomes, and where the sensitivity tends to be stronger.

There may be several dose-event gradient estimates for a study because non-monotonically increasing doses tend to be used.
E.g. 10mg per day is not necessarily greater than or less than 50mg per week because the delivered quantity has increased but the frequency has decreased.
Whether the combined effects increase or decrease toxicity risk will be determined by how the body deals with the drug.
An analagous problem is commonly seen in dose-finding trials of two agents.
Our pragmatic solution is to analyse subsets of doses tha *are* unambiguously orderable.
This yields multiple estimates per study in some cases.
Collapse the estimates by simple averageing:

```{r}
dlt_models %>% 
  group_by(Study) %>% 
  summarise(dlt_dose_grad = mean(dose_grad)) %>% 
  inner_join(
    obj_resp_models %>% 
      group_by(Study) %>% 
      summarise(obj_resp_dose_grad = mean(dose_grad)),
    by = 'Study'
  ) -> paired_dose_sensitivities
```

We only have data here when a trial reports both DLT and objective response.
We may answer the question, how commonly is the dose-sensitivity stronger in DLT than response?
```{r}
paired_dose_sensitivities %>% 
  summarise(mean(dlt_dose_grad > obj_resp_dose_grad))
```

Further evidence that toxicity is more sensitive to dose increases.
Let us visualise the fitted parameters.

```{r}
paired_dose_sensitivities %>% 
  ggplot(aes(x = dlt_dose_grad, y = obj_resp_dose_grad)) + 
  geom_point() + 
  geom_abline(col = 'red', linetype = 'dashed') + 
  labs(x = 'Sensitivity of DLT to dose',
       y = 'Sensitivity of objective response to dose')
```

We see that some implausible sensitivities are fit, with absolute values commonly exceeding 10.
However, there is general evidence that sensitivities are greater for DLT - most points lay below the 45-degree line.

We could zoom in on the absolute values fit between [0, 2].
Perhaps we are more confident in these values:

```{r}
paired_dose_sensitivities %>% 
  left_join(studies, by = 'Study') %>% 
  filter(abs(dlt_dose_grad) <= 2 & abs(obj_resp_dose_grad) <= 2) %>% 
  ggplot(aes(x = dlt_dose_grad, y = obj_resp_dose_grad)) + 
  geom_point(aes(col = DoseVaryingTreatmentType)) + 
  geom_abline(col = 'red', linetype = 'dashed') + 
  labs(x = 'Sensitivity of DLT to dose increases',
       y = 'Sensitivity of objective response to dose increases')
```

Once again, the trend is for values to lay below the $y=x$ line, suggesting that DLT is more sensitive than efficacy to increases in dose.

The implication of this message is that when we escalate dose in dose-finding trials, we are likely embracing higher chances of toxicity faster than we are achieving increases in the probability of response.

These inferences are tempered somewhat by the small individual sample sizes and the computational problems.
We consider an alternative analysis by Bayesian hierarchical models.

# Bayesian hierarchical models
Bayesian hierarchical models have the attraction of partially pooling estimates of intercepts and gradients by assuming exchangeability within a distribution of group-level effects.
We fit models for DLT and response outcomes that assume:

* each series takes its own intercept and dose sensitivity;
* each treatment type also confers its own intercept and dose sensitivity;
* effects are additive.

Thus, this model form allows us to estimate the extent to which dose-event sensitivity is positive for each drug type.

Fitting the models using `brms`:

```{r, warning=FALSE, results='hide'}
library(brms)

dlt_mod <- brm(Events | trials(n) ~ (1 + DoseLevel | AnalysisSeriesId) + 
                 (1 + DoseLevel | DoseVaryingTreatmentType), 
               data = dlt %>% left_join(studies, by = 'Study'), 
               family = binomial('logit'), 
               control = list(adapt_delta = 0.95), cores = 4, seed = 123,
               file = 'dlt_mod')

obj_resp_mod <- brm(Events | trials(n) ~ (1 + DoseLevel | AnalysisSeriesId) + 
                      (1 + DoseLevel | DoseVaryingTreatmentType), 
                    data = obj_resp %>% 
                      left_join(studies, by = 'Study') %>% 
                      filter(n > 0), 
                    family = binomial('logit'), 
                    control = list(adapt_delta = 0.99), cores = 4, seed = 123,
                    file = 'obj_resp_mod')
```

```{r}
summary(dlt_mod)
```

```{r}
summary(obj_resp_mod)
```

The models fit with no warnings of divergences or violations of maximum tree-depth.
The effective sample sizes look reasonable.

We use `tidybayes` to add posterior fitted draws and infer the implied dose-DLT curve from each series.

```{r, warning=FALSE}
dlt %>% 
  left_join(studies, by = 'Study') %>% 
  filter(DoseVaryingTreatmentType %in% c('Chemotherapy', 
                                         'Radiotherapy', 
                                         'Monoclonal Antibody', 
                                         'Inhibitor')) %>% 
  add_fitted_draws(dlt_mod) %>%
  mutate(EstProbEvent = .value / n) %>% 
  summarise(EstProbEvent = mean(EstProbEvent)) %>% 
  ggplot(aes(x = DoseLevel, y = EstProbEvent, group = AnalysisSeriesId, 
             col = DoseVaryingTreatmentType)) + 
  geom_line(size = 1) + 
  facet_wrap(~ DoseVaryingTreatmentType, ncol = 2) + 
  xlim(-5, 5) +
  labs(title = 'Fitted dose-DLT curves', col = 'Drug type', y = 'Prob(DLT)') + 
  theme(legend.position = 'none')
```

There are other treatment types but we visualise just these four for clarity.
Certainly these are the staples of modern oncology treatment.
We see that dose senstivity could safely be taken to be positive under chemo, inhibitors, and radiotherapy.
The relationship is rather less prounounced for mononclonal antibodies.
This is confirmed by visualising the posterior distribution of the dose-sensitivitites that the model attributes to treatment type, i.e. distinct from effects attributable to individual heterogeneity in series. 

```{r, message=FALSE}
dlt_mod %>% 
  spread_draws(r_DoseVaryingTreatmentType[TmtType,term]) %>%
  filter(TmtType %in% c('Radiotherapy', 'Monoclonal.Antibody', 'Chemotherapy', 
                        'Inhibitor', 'Immunomodulatory.drug')) %>% 
  mutate(TmtTypeO = factor(TmtType, levels = c('Monoclonal.Antibody',
                                               'Immunomodulatory.drug',
                                               'Radiotherapy',
                                               'Inhibitor',
                                               'Chemotherapy'), 
                           ordered = TRUE)) %>%
  filter(term == 'DoseLevel') %>% 
  ggplot(aes(y = TmtTypeO, group = TmtTypeO, fill = TmtTypeO, 
             x = r_DoseVaryingTreatmentType)) + 
  geom_density_ridges() + 
  labs(x = 'Gradient with respect to dose', y = 'Treatment type',
       title = 'Distribution of sensitivity of DLT to dose increases') +
  theme(legend.position = 'none') + 
  xlim(-1.5, 2.5)
```
This plot creates a nice cascade.
We might not be surprised that chemo demonstrates the strongest evidence that higher doses are more likely to yield DLT, given that chemo treatments are cytotoxic.
Notably, inhibitor drugs, also commonly called molecularly targeted agents, exhibit strong evidence of DLT risk increasing in dose.
That relationship is much less pronounced for the relatively low numer of antibody drugs in this dataset.

We create the same visualisation for the objective response data.

```{r}
obj_resp %>% 
  filter(n > 0) %>% 
  left_join(studies, by = 'Study') %>% 
  filter(DoseVaryingTreatmentType %in% c('Chemotherapy', 
                                         'Radiotherapy', 
                                         'Monoclonal Antibody', 
                                         'Inhibitor')) %>% 
  add_fitted_draws(obj_resp_mod) %>%
  mutate(EstProbEvent = .value / n) %>% 
  summarise(EstProbEvent = mean(EstProbEvent)) %>% 
  ggplot(aes(x = DoseLevel, y = EstProbEvent, group = AnalysisSeriesId, 
             col = DoseVaryingTreatmentType)) + 
  geom_line(size = 1) + 
  facet_wrap(~ DoseVaryingTreatmentType, ncol = 2) + 
  xlim(-5, 5) +
  labs(title = 'Fitted dose-DLT curves', col = 'Drug type', 
       y = 'Prob(ObjectiveResponse)') + 
  theme(legend.position = 'none')
```

Here the relationships with dose are much less pronounced.
The distributions of the treatment-specific dose effects all straddle zero:

```{r, warning=FALSE, message=FALSE}
obj_resp_mod %>% 
  spread_draws(r_DoseVaryingTreatmentType[TmtType,term]) %>%
  filter(TmtType %in% c('Radiotherapy', 'Monoclonal.Antibody', 'Chemotherapy', 
                        'Inhibitor', 'Immunomodulatory.drug')) %>% 
  mutate(TmtTypeO = factor(TmtType, levels = c('Monoclonal.Antibody',
                                               'Immunomodulatory.drug',
                                               'Radiotherapy',
                                               'Inhibitor',
                                               'Chemotherapy'), 
                           ordered = TRUE)) %>%
  filter(term == 'DoseLevel') %>% 
  ggplot(aes(y = TmtTypeO, group = TmtTypeO, fill = TmtTypeO, 
             x = r_DoseVaryingTreatmentType)) + 
  geom_density_ridges() + 
  labs(x = 'Gradient with respect to dose', y = 'Treatment type',
       title = 'Distribution of sensitivity of response to dose increases') +
  theme(legend.position = 'none') + 
  xlim(-0.5, 1)
```

The Bayesian hierarchical analysis has echoed the inferences from the series-specific models presented above, and provided some additional insights attributable to individual treatments.

There is strong evidence that the risks of DLT increase in dose in many treatments like chemo, inhibitors, and radiotherapy.
The relationship is rather weaker for monoclonal antibodies.
There is much less evidence to support the assumption that the chances of objective response are increased on average by increasing dose.

Thus in dose-finding trials, the data suggest that escalations have tended to coerce greater toxicity risk without commensurate increases in the chances of response.
Dose-effects for response cannot be ruled out based on this analysis.
However, the uncertainty surrounding their presence suggests that trialists should avoid relying unduly on establishing a maximum tolerable dose and instead focus on identifying the dose that offers the most attractive trade-off between the competing outcomes of DLT and response.
Use of so-called seamless phase I/II designs like EffTox (Thall & Cook, 2004), Wages & Tait (2015), TriCRM (Zhang et al, 20??)  and BiCRM (Braun, 20??) could help achieve this goal.

