---
title: "Checks on the Bayesian Hierarchical Models"
author: "Kristian Brock"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
---

```{r}
# This script in the root of the project loads the datasets
source('../../../Load.R')

library(brms)
library(ggplot2)
```


## DLT Model
```{r}
dlt_mod <- readRDS('dlt_mod.rds')
```

```{r}
dlt_mod
```

Rhats and Eff.Samples look fine. 
No divergent transitions or violations of max tree-depth are reported.

```{r}
pp_check(dlt_mod)
```

The x-axis is the sampled number of DLTs, conditional on the number of attempts, using all of the DLT series. 
Posterior samples are a faithful representation of the observed data.

Distribution of series-level intercepts:
```{r}
qplot(ranef(dlt_mod)$AnalysisSeriesId[, , 'Intercept'][,'Estimate'], 
      geom = 'histogram')
```

Normal enough.

Distribution of series-level gradients for DoseLevel:
```{r}
qplot(ranef(dlt_mod)$AnalysisSeriesId[, , 'DoseLevel'][,'Estimate'], 
      geom = 'histogram')
```

Bit of a negative skew but not alarmingly non-normal.

Distribution of series-level intercepts:
```{r}
qplot(ranef(dlt_mod)$DoseVaryingTreatmentType[, , 'Intercept'][,'Estimate'], 
      geom = 'histogram')
```

There are not many of these.
There appears to be some central tendency and symmetry.
Normal is not a poor working model.

Distribution of series-level gradients for DoseLevel:
```{r}
qplot(ranef(dlt_mod)$DoseVaryingTreatmentType[, , 'DoseLevel'][,'Estimate'], 
      geom = 'histogram')
```

Now these do look kinda weird.
Again, small numbers here.

## Objective Response Model
```{r}
obj_resp_mod <- readRDS('obj_resp_mod.rds')
```

```{r}
obj_resp_mod
```

Rhats and Eff.Samples look fine. 
No divergent transitions or violations of max tree-depth are reported.

```{r}
pp_check(obj_resp_mod)
```

The x-axis is the sampled number of responses, conditional on the number of attempts, using all of the response series. 
Posterior samples are a faithful representation of the observed data.

Distribution of series-level intercepts:
```{r}
qplot(ranef(obj_resp_mod)$AnalysisSeriesId[, , 'Intercept'][,'Estimate'], 
      geom = 'histogram')
```

Normal enough.

Distribution of series-level gradients for DoseLevel:
```{r}
qplot(ranef(obj_resp_mod)$AnalysisSeriesId[, , 'DoseLevel'][,'Estimate'], 
      geom = 'histogram')
```

Long tails but not alarmingly non-normal.

Distribution of series-level intercepts:
```{r}
qplot(ranef(obj_resp_mod)$DoseVaryingTreatmentType[, , 'Intercept'][,'Estimate'], 
      geom = 'histogram')
```

There are not many of these.
There appears to be some central tendency and symmetry.
Normal is not a poor working model.

Distribution of series-level gradients for DoseLevel:
```{r}
qplot(ranef(obj_resp_mod)$DoseVaryingTreatmentType[, , 'DoseLevel'][,'Estimate'], 
      geom = 'histogram')
```

Again, small numbers here but apparent long tails.

