---
title: "ModelInferences"
author: "Kristian Brock"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
---

Modelling of the event rates in the observed series is presented in DLTSeriesPresentation and ObjRespSeriesPresentation.
Inferences beyond those series-specific insights are explored here.

```{r}
# # This script in the root of the project loads the datasets
# source('../../../Load.R')

library(brms)
library(tidybayes)
library(ggplot2)
library(ggridges)

# Load DLT model
dlt_mod <- readRDS('dlt_mod.rds')

# Load objective response model
obj_resp_mod <- readRDS('obj_resp_mod.rds')
```

## DLT Model

Mean or baseline or expected event rate:
```{r, message=FALSE, warning=FALSE}
dlt_mod %>% 
  spread_draws(r_DoseVaryingTreatmentType[TmtType,term]) %>%
  filter(TmtType %in% c('Radiotherapy', 'Monoclonal.Antibody', 'Chemotherapy', 
                        'Inhibitor')) %>% 
  mutate(TmtTypeO = factor(TmtType, levels = c('Monoclonal.Antibody',
                                               'Radiotherapy',
                                               'Inhibitor',
                                               'Chemotherapy'), 
                           ordered = TRUE)) %>%
  filter(term == 'Intercept') %>% 
  ggplot(aes(y = TmtTypeO, group = TmtTypeO, fill = TmtTypeO, 
             x = r_DoseVaryingTreatmentType)) + 
  geom_density_ridges() + 
  labs(x = 'Adjustment to log(odds) of DLT compared to entire dataset', 
       y = 'Treatment type',
       title = 'Distributions of baseline rates of DLT') +
  theme(legend.position = 'none') + 
  xlim(-1.5, 2.5)
```
After adjusting for the heterogeneity seen in the individual series, it appears that chemo is probably the therapy type with the greatest average risk of DLT.
However, with inhibitors and radiotherapy, the risk of DLT is almost the same.
The similarity in baseline rates may suggest that DLT is chosen mindful of therapy to be something that is likely to occur in a set proportion of patients.
However, this would require admirable forecasting skills from trialists.
Notably, incidence of DLT is less common in monoclonal antibodies.


Sensitivity of event to increases in dose
```{r}
dlt_mod %>% 
  spread_draws(r_DoseVaryingTreatmentType[TmtType,term]) %>%
  filter(TmtType %in% c('Radiotherapy', 'Monoclonal.Antibody', 'Chemotherapy', 
                        'Inhibitor')) %>% 
  mutate(TmtTypeO = factor(TmtType, levels = c('Monoclonal.Antibody',
                                               'Radiotherapy',
                                               'Inhibitor',
                                               'Chemotherapy'), 
                           ordered = TRUE)) %>%
  filter(term == 'DoseLevel') %>% 
  ggplot(aes(y = TmtTypeO, group = TmtTypeO, fill = TmtTypeO, 
             x = r_DoseVaryingTreatmentType)) + 
  geom_density_ridges() + 
  labs(x = 'Adjustment to gradient with respect to dose', 
       y = 'Treatment type',
       title = 'Distribution of sensitivity of DLT to dose increases') +
  theme(legend.position = 'none') + 
  xlim(-1.5, 2.5)
```
With chemotherapies, inhibitors, and radiotherapy, the relationship of DLT risk with dose is concincingly positive. 
That is, greater doses of these treatments are associated with greater risk of DLT.
The message is rather more opaque with antibodies.
The aggregate evidence supports a positive relationship, but the strenght of the relationship is rather weaker than the other drug classes and could plausbily be very small.


All treatment types:
```{r, warning=FALSE, message=FALSE}
dlt_mod %>% 
  spread_draws(r_DoseVaryingTreatmentType[TmtType,term]) %>%
  filter(term == 'DoseLevel') %>% 
  ggplot(aes(y = TmtType, group = TmtType, fill = TmtType, 
             x = r_DoseVaryingTreatmentType)) + 
  geom_density_ridges() + 
  labs(x = 'Adjustment to gradient with respect to dose', 
       y = 'Treatment type',
       title = 'Distribution of sensitivity of DLT to dose increases') +
  theme(legend.position = 'none') + 
  xlim(-1.5, 2.5)
```
There is more positive mass than negative mass, suggesting that the opposite must be seen in the AnalysisSeriesId-level analyses?

```{r, warning=FALSE, message=FALSE, fig.height = 20, fig.width = 8}
dlt_mod %>% 
  spread_draws(r_AnalysisSeriesId[AnalysisSeriesId,term]) %>%
  filter(term == 'DoseLevel') %>% 
  ggplot(aes(y = AnalysisSeriesId, group = AnalysisSeriesId, 
             fill = AnalysisSeriesId, x = r_AnalysisSeriesId)) + 
  geom_density_ridges() + 
  labs(x = 'Adjustment to gradient with respect to dose', 
       title = 'Distribution of sensitivity of DLT to dose increases') +
  theme(legend.position = 'none') + 
  xlim(-1.5, 2.5)
```

Woof. How would you know?


## Objective Response Model

Mean or baseline or expected event rate:
```{r, message=FALSE, warning=FALSE}
obj_resp_mod %>% 
  spread_draws(r_DoseVaryingTreatmentType[TmtType,term]) %>%
  filter(TmtType %in% c('Radiotherapy', 'Monoclonal.Antibody', 'Chemotherapy', 
                        'Inhibitor')) %>% 
  mutate(TmtTypeO = factor(TmtType, levels = c('Monoclonal.Antibody',
                                               'Radiotherapy',
                                               'Inhibitor',
                                               'Chemotherapy'), 
                           ordered = TRUE)) %>%
  filter(term == 'Intercept') %>% 
  ggplot(aes(y = TmtTypeO, group = TmtTypeO, fill = TmtTypeO, 
             x = r_DoseVaryingTreatmentType)) + 
  geom_density_ridges() + 
  labs(x = 'Adjustment to log(odds) of Objective Response compared to entire dataset', 
       y = 'Treatment type',
       title = 'Distributions of baseline rates of Objective Response') +
  theme(legend.position = 'none') + 
  xlim(-2, 3)
```
Chemo is associated with strongest average response inducing effect.
Inhibitors the least.


Sensitivity of event to increases in dose
```{r}
obj_resp_mod %>% 
  spread_draws(r_DoseVaryingTreatmentType[TmtType,term]) %>%
  filter(TmtType %in% c('Radiotherapy', 'Monoclonal.Antibody', 'Chemotherapy', 
                        'Inhibitor')) %>% 
  mutate(TmtTypeO = factor(TmtType, levels = c('Monoclonal.Antibody',
                                               'Radiotherapy',
                                               'Inhibitor',
                                               'Chemotherapy'), 
                           ordered = TRUE)) %>%
  filter(term == 'DoseLevel') %>% 
  ggplot(aes(y = TmtTypeO, group = TmtTypeO, fill = TmtTypeO, 
             x = r_DoseVaryingTreatmentType)) + 
  geom_density_ridges() + 
  labs(x = 'Adjustment to gradient with respect to dose', 
       y = 'Treatment type',
       title = 'Distribution of sensitivity of Objective Response to dose increases') +
  theme(legend.position = 'none') + 
  xlim(-1, 1)
```
Every mode is at zero.
Some tails are looong.