---
title: "Run DLT and objective-response models"
author: "Kristian Brock"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_notebook
---

This is done in a workbook to take advantage of the automatic setting of working directory.

```{r}
# This script in the root of the project loads the datasets
source('../../../Load.R')

library(rstanemax)
library(brms)
library(ggplot2)
library(patchwork)

emax <- function(d, E0, EMax, ED50, lambda) {
  E0 + (EMax * d^lambda) / (ED50 + d^lambda)
}

fit_emax_dlt <- function(analysis_series_id, ...) {
  dlt %>% 
    filter(AnalysisSeriesId == analysis_series_id) %>% 
    mutate(
      ProbEvent = ifelse(ProbEvent == 0, ProbEvent + 0.001, ProbEvent),
      ProbEvent = ifelse(ProbEvent == 1, ProbEvent - 0.001, ProbEvent),
      DoseLevelN = 1:nrow(.),
      LogOddsEvent = logit_scaled(ProbEvent)
    ) -> model_dat
  
  fit <- stan_emax(LogOddsEvent ~ DoseLevelN, gamma.fix = NULL, 
                   data = model_dat, 
                   priors = list(gamma = c(0, 1)),  # Default is N(0, 3)
                   control = list(adapt_delta = 0.999), iter = 5000, ...)
  
  list(data = model_dat, fit = fit)
}

fit_emax_obj_resp <- function(analysis_series_id, ...) {
  obj_resp %>% 
    filter(AnalysisSeriesId == analysis_series_id) %>% 
    mutate(
      ProbEvent = ifelse(ProbEvent == 0, ProbEvent + 0.001, ProbEvent),
      ProbEvent = ifelse(ProbEvent == 1, ProbEvent - 0.001, ProbEvent),
      DoseLevelN = 1:nrow(.),
      LogOddsEvent = logit_scaled(ProbEvent)
    ) -> model_dat
  
  fit <- stan_emax(LogOddsEvent ~ DoseLevelN, gamma.fix = NULL, 
                   data = model_dat, 
                   priors = list(gamma = c(0, 1)),  # Default is N(0, 3)
                   control = list(adapt_delta = 0.999), iter = 5000, ...)
  
  list(data = model_dat, fit = fit)
}

get_dlt_label <- function(series_id) {
  dlt %>% 
    left_join(studies, by = 'Study') %>% 
    filter(AnalysisSeriesId == series_id) %>% 
    select(Study, DoseVaryingTreatment, DoseVaryingTreatmentType, 
           PatientGroup) %>% 
    mutate(Label = paste0(DoseVaryingTreatment, ' (', 
                          DoseVaryingTreatmentType, ') in ', 
                          stringr::str_to_title(PatientGroup), ' [',
                          Study, ']'
    )) %>% 
    select(Label) %>% distinct() %>% .[[1]] -> label
  
  label
}

get_obj_resp_label <- function(series_id) {
  obj_resp %>% 
    left_join(studies, by = 'Study') %>% 
    filter(AnalysisSeriesId == series_id) %>% 
    select(Study, DoseVaryingTreatment, DoseVaryingTreatmentType, 
           PatientGroup) %>% 
    mutate(Label = paste0(DoseVaryingTreatment, ' (', 
                          DoseVaryingTreatmentType, ') in ', 
                          stringr::str_to_title(PatientGroup), ' [',
                          Study, ']'
    )) %>% 
    select(Label) %>% distinct() %>% .[[1]] -> label
  
  label
}

plot_lines <- function(x, use_rows = 1000) {
  preds <-  posterior_predict(x$fit, returnType = "tibble")

  preds %>% 
    head(use_rows) %>% 
    mutate(logodds_event = emax(exposure, E0 = e0, EMax = emax, ED50 = ec50, 
                                lambda = gamma),
           prob_event = inv_logit_scaled(logodds_event)) %>% 
    ggplot(aes(x = exposure, y = prob_event)) +
    geom_line(aes(group = mcmcid), alpha = 0.1, col=  'darkgreen') + 
    ylim(0, 1) + 
    labs(x = 'Dose-level', y = '') + 
    geom_point(data = x$data, aes(x = DoseLevelN, y = ProbEvent, 
                                     size = n), col = 'black')
}

plot_mean <- function(x) {
  preds <-  posterior_predict(x$fit, returnType = "tibble")

  preds %>% 
    mutate(logodds_event = emax(exposure, E0 = e0, EMax = emax, ED50 = ec50, 
                                lambda = gamma),
           prob_event = inv_logit_scaled(logodds_event)) %>% 
    group_by(exposure) %>% 
    summarise(prob_event = mean(prob_event)) %>% 
    ggplot(aes(x = exposure, y = prob_event)) +
    geom_line(col=  'darkgreen') + 
    ylim(0, 1) + 
    labs(x = 'Dose-level', y = '') + 
    geom_point(data = x$data, aes(x = DoseLevelN, y = ProbEvent, 
                                     size = n), col = 'black')
}
```


# DLT
```{r}
set.seed(68764521)

dlt_models <- list()
analysis_series_ids <- dlt %>% 
  distinct(AnalysisSeriesId) %>% .[[1]]

for(analysis_series_id in analysis_series_ids) {
  
  print(analysis_series_id)
  mod <- fit_emax_dlt(analysis_series_id, refresh = 0)
  dlt_models[[as.character(analysis_series_id)]] <- mod
}

length(dlt_models)
```

```{r}
format(object.size(dlt_models), units = 'MB')
```

```{r}
library(purrr)
```

Convert from list to tibble for simpler analysis.
```{r}
tibble(
  data = dlt_models %>% map('data'),
  fit = dlt_models %>% map('fit')
) -> dlt_series_summary

dlt_series_summary %>% 
  mutate(
    Study = data %>% map_chr(function(x) x %>% 
                               distinct(Study) %>% 
                               .[[1]]),
    AnalysisSeriesId = data %>% map_dbl(function(x) x %>% 
                                          distinct(AnalysisSeriesId) %>% 
                                          .[[1]]),
    Events = data %>% map_dbl(function(x) x %>%
                                summarise(sum(Events)) %>%
                                .[[1]]),
    NPostSamples = fit %>% map_int(function(x) x$stanfit %>%
                                     as.data.frame %>%
                                     nrow)
  ) -> dlt_series_summary

dlt_series_summary %>% 
  select(-data, -fit) %>% 
  head
```

```{r, rows.print = 100}
dlt_series_summary %>% 
  select(-data, -fit) %>% 
  filter(NPostSamples == 0)
```

Series 91 failed despite 7 events.
Series 597 failed despite 6 events.
Otherwise, fitting only fails when Events == 0.

Did fitting ever succeed when Events == 0?
```{r}
dlt_series_summary %>% 
  select(-data, -fit) %>% 
  filter(Events == 0 & NPostSamples > 0)
```
Of course not!

```{r}
format(object.size(dlt_series_summary), units = 'MB')
```
Storing expensive objects in two places now.
Just be mindful of memory. 

Take a look at some:
```{r, fig.width=10}
id <- '3'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '6'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '21'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '32'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '39'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '86'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r}
dlt %>% filter(Study == 'Sharma2013_1')
```

```{r, fig.width=10}
id <- '100'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r}
dlt %>% filter(Study == 'Sharma2013_2')
```

```{r, fig.width=10}
id <- '105'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r}
dlt_series_summary %>% filter(Study == 'Saji2008_1')
```

```{r, fig.width=10}
id <- '325'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

```{r}
dlt_series_summary %>% filter(Study == 'Kim2009_1')
```

```{r, fig.width=10}
id <- '496'
label <- get_dlt_label(id)
p1 <- plot_lines(dlt_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(dlt_models[[id]])
p1 + p2
```

# Obj Resp
```{r}
set.seed(68764521)

obj_resp_models <- list()
analysis_series_ids <- obj_resp %>% 
  distinct(AnalysisSeriesId) %>% .[[1]]

for(analysis_series_id in analysis_series_ids) {
  
  print(analysis_series_id)
  mod <- fit_emax_obj_resp(analysis_series_id, refresh = 0)
  obj_resp_models[[as.character(analysis_series_id)]] <- mod
}

length(obj_resp_models)
```

```{r}
format(object.size(obj_resp_models), units = 'MB')
```

```{r}
tibble(
  data = obj_resp_models %>% map('data'),
  fit = obj_resp_models %>% map('fit')
) -> obj_resp_series_summary

obj_resp_series_summary %>% 
  mutate(
    Study = data %>% map_chr(function(x) x %>% 
                               distinct(Study) %>% 
                               .[[1]]),
    AnalysisSeriesId = data %>% map_dbl(function(x) x %>% 
                                          distinct(AnalysisSeriesId) %>% 
                                          .[[1]]),
    Events = data %>% map_dbl(function(x) x %>%
                                summarise(sum(Events)) %>%
                                .[[1]]),
    NPostSamples = fit %>% map_int(function(x) x$stanfit %>%
                                     as.data.frame %>%
                                     nrow)
  ) -> obj_resp_series_summary

obj_resp_series_summary %>% 
  select(-data, -fit) %>% 
  head
```

```{r, rows.print = 100}
obj_resp_series_summary %>% 
  select(-data, -fit) %>% 
  filter(NPostSamples == 0)
```

Series 98 failed despite 4 events.
Series 157 failed despite 19 events.
Otherwise, fitting only fails when Events == 0.

Did fitting ever succeed when Events == 0?
```{r}
obj_resp_series_summary %>% 
  select(-data, -fit) %>% 
  filter(Events == 0 & NPostSamples > 0)
```
Of course not!

Take a look at some:

```{r}
obj_resp_series_summary %>% 
  select(-data, -fit) %>% 
  filter(NPostSamples > 0) %>% head
```

```{r, fig.width=10}
id <- '28'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '37'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '46'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '59'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '66'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```

```{r, fig.width=10}
id <- '75'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```

```{r}
obj_resp_series_summary %>% filter(Study == 'Saji2008_1')
```

```{r, fig.width=10}
id <- '326'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```

```{r}
obj_resp_series_summary %>% filter(Study == 'Kim2009_1')
```

```{r, fig.width=10}
id <- '501'
label <- get_obj_resp_label(id)
p1 <- plot_lines(obj_resp_models[[id]], use_rows = 1000) + 
  labs(title = label)
p2 <- plot_mean(obj_resp_models[[id]])
p1 + p2
```


# Fitted Series
```{r}
library(tidyr)

seed <- 98592

set.seed(seed)
dlt_series_summary %>% 
  filter(NPostSamples > 0) %>% 
  mutate(preds = fit %>% map(posterior_predict, returnType = "tibble")) %>% 
  select(-data, -fit) %>% 
  unnest() %>% 
  mutate(logodds_event = emax(exposure, E0 = e0, EMax = emax, ED50 = ec50, 
                              lambda = gamma),
         prob_event = inv_logit_scaled(logodds_event)) %>% 
  rename(DoseLevelN = exposure) %>% 
  group_by(Study, AnalysisSeriesId, DoseLevelN) %>% 
  summarise(prob_event = mean(prob_event)) %>% 
  left_join(studies, by = 'Study') %>% 
  left_join(manuscripts, by = 'Manuscript') %>% 
  left_join(dlt, by = c('AnalysisSeriesId' = 'AnalysisSeriesId', 
                        'DoseLevelN' = 'DoseLevelN')) -> fitted_emax_dlt_series


set.seed(seed)
obj_resp_series_summary %>% 
  filter(NPostSamples > 0) %>% 
  mutate(preds = fit %>% map(posterior_predict, returnType = "tibble")) %>% 
  select(-data, -fit) %>% 
  unnest() %>% 
  mutate(logodds_event = emax(exposure, E0 = e0, EMax = emax, ED50 = ec50, 
                              lambda = gamma),
         prob_event = inv_logit_scaled(logodds_event)) %>% 
  rename(DoseLevelN = exposure) %>% 
  group_by(Study, AnalysisSeriesId, DoseLevelN) %>% 
  summarise(prob_event = mean(prob_event)) %>% 
  left_join(studies, by = 'Study') %>% 
  left_join(manuscripts, by = 'Manuscript') %>% 
  left_join(obj_resp, by = c('AnalysisSeriesId' = 'AnalysisSeriesId', 
                             'DoseLevelN' = 'DoseLevelN')) -> 
  fitted_emax_obj_resp_series
```

